---
title: "Covid91 vaccine study Final2023"
author: "Anthony Falconio"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


$H_0$ : The standard hypothesis states that there are no statistical differences in the data, which implies that the vaccine and the placebo are equally effective

$H_A$ : The alternate hypothesis states that there is a significant difference in the data, which implies that the vaccine and placebo are NOT equally effective.

##Methods

Upon analyzing the gathered data, we observe that this analysis involves a categorical variable explained by another categorical variable. To facilitate the isolation and testing of each group, distinct data variables will be generated for individual groups. The candP helper file will be employed to utilize R Studio's built-in bar charts for visually representing the data. Subsequently, xtabs will be utilized to compute the pertinent percentages for the dataset. In this instance, as we are exclusively comparing two types of data, there is no need to employ loops for assessing multiple variables in the Fisher test or executing a FisherXY.

# Results

## Males

### Graphical Results

```{r}
barchartGC(~infected + treatment,data=male)
barchartGC(~infected + treatment,data=male, type="percent")
```

Displayed above is a bar chart illustrating the visual representation of outcomes exclusively for males derived from the data. The comparison focuses on the incidence of contracting Covid-19 based on vaccination status. The left side depicts individuals who contracted Covid-19, while the right side illustrates males who did not.

The visual data clearly indicates that the percentage of individuals in the placebo group who contracted Covid-19 is more than double compared to those who did not contract the virus, and this contrast is noticeably less pronounced among those who remained Covid-free. Considering the sample size, the likelihood of this discrepancy occurring solely by chance is improbable.

### Numerical Results

```{r}
tableMale <- xtabs(~infected + treatment, data=male)
rowPerc(tableMale)
colPerc(tableMale)
```

The numerical findings corroborate the insights conveyed by the graphical results. Specifically, the placebo group exhibited a Covid contraction rate of approximately 70%, while the vaccinated group experienced a Covid contraction rate of 30%. These figures underscore the overall effectiveness of the vaccine in this testing scenario.

###Inferential Results

```{r}
chisq.test(tableMale)
chisqtestGC(tableMale)
maleFisher <- fisher.test(tableMale)
maleFisher
```

The results of the Chi-Squared test reveal a significantly low P-value for the analyzed data, far below the conventional threshold of 0.05. In fact, it stands at $1.098\times10^{-13}$, an exceedingly small value indicative of a substantial difference in the data. Furthermore, the Chi-Squared statistic registers at 55, considerably surpassing the one degree of freedom associated with this dataset.

Remarkably, the odds of an unvaccinated man contracting Covid-19 are 2.2 times higher than those of a vaccinated man falling to the virus.

### Confidence Interval
```{r}
oddsMale = maleFisher$estimate/(maleFisher$estimate+1)*100
print(oddsMale)
maleLow=(maleFisher$conf.int[1]/(maleFisher$conf.int[1]+1))*100
maleHigh=(maleFisher$conf.int[2]/(maleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Males: ",maleLow," to ",maleHigh)
```

Observing the data, it is evident that the confidence interval exceeds 50%, surpassing the standard vaccine threshold of 60%. This confidence interval is deemed more than sufficient to affirm the effectiveness of the vaccine within this specific group.

## Females

## Graphical Results

```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")
```

Displayed above is a bar chart presenting the graphical outcomes exclusively for females gathered from the dataset. The comparison focuses on the incidence of contracting Covid-19 based on vaccination status, with the left side representing women who contracted Covid-19 and the right side depicting women who remained free of the virus.

It is evident that approximately one-third more of the percentage of individuals in the placebo group contracted Covid-19, while those who did not contract the virus exhibit closely aligned proportions.

## Numerical Results

```{r}
tableFemale <- xtabs(~infected + treatment, data=female)
rowPerc(tableFemale)
colPerc(tableFemale)
```

The findings align with the insights derived from the graphical representation of the data. The placebo group exhibited a Covid contraction rate of around 60%, while the vaccinated group demonstrated a 40% incidence. It is evident that the vaccine generally proves effective in this testing scenario, although its efficacy appears to be lower than observed among male participants.

## Inferential Results

```{r}
chisq.test(tableFemale)
chisqtestGC(tableFemale)
femaleFisher <- fisher.test(tableFemale)
femaleFisher
```


The results of the Chi-Squared test indicate a P value for the tested data that is below the commonly accepted threshold of 0.05. In fact, it is $2.732\times10^{-4}$, signifying a significantly small value and suggesting a notable difference in the data. The Chi-Squared statistic, measuring 13, exceeds the degrees of freedom (one) derived from this dataset.

Furthermore, the analysis reveals that the odds of an unvaccinated woman contracting Covid-19 are 1.5 times higher than those of a woman with the vaccine.

### Confidence Interval

```{r}
oddsFemale = femaleFisher$estimate/(femaleFisher$estimate +1)*100
print(oddsFemale)
femaleLow=(femaleFisher$conf.int[1]/(femaleFisher$conf.int[1]+1))*100
femaleHigh=(femaleFisher$conf.int[2]/(femaleFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for Females: ",femaleLow," to ",femaleHigh)
```

Once more, it is evident that the confidence interval surpasses the 50% mark and encompasses the 60% threshold. While its efficacy is lower compared to the male group, the vaccine remains effective in this context.

## LGBTQ

### Graphical Results

```{r}
barchartGC(~infected + treatment,data=lgbtq)
barchartGC(~infected + treatment,data=lgbtq, type="percent")
```

Presented above is a bar chart that juxtaposes the visual outcomes exclusively for individuals who are drug users, sourced from the data. This comparison examines the rates of contracting Covid-19 based on vaccination status, with the left side representing individuals who contracted the virus and the right side illustrating those who did not.

Once more, the discernible pattern reveals a significant increase in the number of individuals from the vaccinated group who contracted Covid-19. Conversely, there is a noteworthy proximity among those who remained free of the virus. Considering the sample size, the probability of this occurrence being purely random is unlikely.

### Numerical Results

```{r}
tableLGBTQ <- xtabs(~infected + treatment, data=lgbtq)
rowPerc(tableLGBTQ)
colPerc(tableLGBTQ)
```

The numerical results reinforce the insights from the visual representation, revealing that the placebo group experienced Covid contraction at a rate of approximately 15%, while the vaccinated group exhibited an 85% incidence of contracting Covid. Contrary to expectations, it becomes evident that the vaccine is ineffective in this testing scenario. In fact, it appears to have an inverse effect, potentially placing this group at a heightened risk of contracting Covid.

### Inferential Results

```{r}
chisq.test(tableLGBTQ)
chisqtestGC(tableLGBTQ)
lgbtqFisher <- fisher.test(tableLGBTQ)
lgbtqFisher
```

The Chi-Squared test reveals a significantly low P value for the examined data, well below the threshold of 0.05; in fact, it stands at $6.668\times10^{-8}$. This minute value suggests a notable difference in the data. The Chi-Squared statistic, registering at 29, surpasses our degrees of freedom, which is one for this dataset. In this group, the likelihood of a vaccinated individual contracting Covid-19 is approximately 1 in 5, highlighting a noteworthy risk.

### Confidence Interval

```{r}
oddsLGBTQ = lgbtqFisher$estimate/(lgbtqFisher$estimate +1)*100
print(oddsLGBTQ)
lgbtqLow=(lgbtqFisher$conf.int[1]/(lgbtqFisher$conf.int[1]+1))*100
lgbtqHigh=(lgbtqFisher$conf.int[2]/(lgbtqFisher$conf.int[2]+1))*100
cat("\n","Condifence interval for LGBTQ: ",lgbtqLow," to ",lgbtqHigh)
```

Observing the confidence interval ranging from 7% to 26%, it is evident that this falls significantly below the 50% threshold. Consequently, we can conclude that this vaccine may not be deemed highly effective.

## Druggies


# Overall Results and Conclusions